---
title: "Coronavirus Time"
author: "Anthony Staines"
date: "13/03/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#Purpose
To prepare models for coronavirus incidence, and short term predictions, in Ireland

```{r setup, include=FALSE}
library(coronavirus) # Coronavirus data
library(propagate) # Errors in non-linear regression models
library(ggrepel) # Nice labels
library(lubridate) # Sane date handling
library(drc)
library(sandwich)
library(lmtest)
library(tidyverse)

data("coronavirus")

knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

```{r load data file}
load('data/CC.Rdata')

```
## Identify the areas and the first date of a recorded confirmed case, and of the first recorded death

```{r Country}
Areas <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(Country.Region) %>%
  summarise(Total.Cases = sum(cases)) %>%
  arrange(desc(Total.Cases), Country.Region)

FirstDates <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(Country.Region) %>%
  filter(cases > 0) %>%
  filter(date == min(date)) %>%
  select(Country.Region, First.Date = date) %>%
  distinct()  %>%
  arrange(First.Date, Country.Region)
  
DeathDates <- CC %>%
  filter(type == 'death') %>%
  group_by(Country.Region) %>%
  filter(cases > 0) %>%
  filter(date == min(date)) %>%
  select(Country.Region, First.Date = date) %>%
  distinct()  %>%
  arrange(First.Date, Country.Region)

```

# Prepare a general data set of confirmed cases

```{r General data set}

Prepared <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(Country.Region, date ) %>%
  summarise(Total.Cases = sum(cases)) %>% # Collapse over Province.State
  right_join(FirstDates) %>%
  arrange(date) %>%
  mutate(Day = seq_along(date)) %>% # Set up days from first case
  filter(Day >= 0) %>%
  arrange(Country.Region, Day) %>%
  mutate(Sum.Cases = cumsum(Total.Cases)) # Cumulative total of cases is Sum.Cases

```

## Prepare an Ireland specific dataset

```{r Ireland specific dataset}
Ireland.Prepared <- CC %>%
  filter(Country.Region == 'Ireland') %>%
  filter(type == 'confirmed') %>%
  group_by(Country.Region, date ) %>%
  summarise(Total.Cases = sum(cases)) %>%
  right_join(FirstDates) %>%
  mutate(Day  = seq_along(date)) %>%
  filter(Day >= 0) %>%
  arrange(Country.Region, Day) %>%
  mutate(Sum.Cases = cumsum(Total.Cases))

```


```{r define prediction frame}
Model.Data <- Prepared %>%
  filter(Country.Region == 'Ireland') %>%
  filter(Sum.Cases > 0) %>% # First Day with 1 case
  mutate(Day.Original = Day) %>%
  mutate(Day = 1 + (Day - min(Day)))

pred.frame=tibble(Day = seq(1 : ( max(Model.Data$Day) + 10))) # 10 days ahead.

```

# Predictions of log(accumulated cases)


## LL model

```{r drc LL}
## LL model

drc.LL.4 <- drm(Sum.Cases ~ Day,
       data =Model.Data, 
                   fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
drc.LL.5 <- drm(Sum.Cases ~ Day,
       data =Model.Data, 
                   fct = LL.5(names = c("Slope", "Lower Limit", "Upper Limit", "ED50", "f")))
summary(drc.LL.4)
AIC(drc.LL.4)
summary(drc.LL.5)
AIC(drc.LL.5)
modelFit(drc.LL.4)
modelFit(drc.LL.5)
plot(drc.LL.4)
plot(drc.LL.5)
coeftest(drc.LL.4, vcov = sandwich)
predict(drc.LL.4,as.data.frame(pred.frame),vcov. = sandwich)
coeftest(drc.LL.5, vcov = sandwich)
predict(drc.LL.5,as.data.frame(pred.frame),vcov. = sandwich)
model.LL.predicted <-
  predict(drc.LL.5,as.data.frame(pred.frame),
        se.fit=TRUE, interval='confidence',
        vcov. = sandwich, od=FALSE, level=0.95) %>%
  as_tibble() %>%
  cbind(pred.frame) %>%
  rename(Prediction.LL = Prediction) %>%
  mutate(Lower.LL = Prediction.LL - 1.96*SE) %>%
  mutate(Upper.LL = Prediction.LL + 1.96*SE) %>%
  select(Day,Prediction.LL,Lower.LL,Upper.LL)

```

## Gompertz model

```{r drc Gompertz model}
##Gompertz model

drc.Gom <- drm(Sum.Cases ~ Day,
       data = Model.Data, 
                   fct = G.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
summary(drc.Gom)

AIC(drc.Gom)
plot(drc.Gom)
modelFit(drc.Gom)
coeftest(drc.Gom, vcov = sandwich)
model.Gom.predicted <-
  predict(drc.Gom,as.data.frame(pred.frame),
        se.fit=TRUE, interval='confidence',
        vcov. = sandwich, od=FALSE, level=0.95) %>%
  as_tibble() %>%
  cbind(pred.frame) %>%
  rename(Prediction.Gom = Prediction) %>%
  mutate(Lower.Gom = Prediction.Gom - 1.96*SE) %>%
  mutate(Upper.Gom = Prediction.Gom + 1.96*SE) %>%
  select(Day,Prediction.Gom,Lower.Gom,Upper.Gom)

```

### Create a file of results

```{r Pull together results}
Results <- Model.Data %>%
  ungroup() %>%
  filter(Country.Region=='Ireland') %>%
  right_join(model.LL.predicted) %>%
  right_join(model.Gom.predicted)
```

## Very quick plot of data

```{r plot1}
ggplot(data=Prepared,aes(x=Day,y=log10(Sum.Cases),colour=Country.Region)) +
  geom_path() +
  geom_point(data=Model.Data,
             aes(x=Day.Original,y=log10(Sum.Cases)),
                 colour='red') +
  theme(legend.position="none")

```

## Plot 1 Ireland against EU  (mostly) 20 cases

```{r}

Prepared.20 <- Prepared %>%
  group_by(Country.Region) %>%
  filter(Sum.Cases > 20) %>%
  mutate(Day2 = seq_along(date))

Prepared.20 <- Prepared.20 %>%
  filter(Country.Region %in% c('New Zealand', 'Hungary', 'France', 'Spain', 'US', 'United Kingdom','Ireland', 'Norway', 'Denmark', 'Finland', 'Germany'))

g <- ggplot(data=Prepared.20,
       aes(x=Day2,y=Sum.Cases,colour=Country.Region)) +
  geom_path() +
  geom_point(data = Prepared.20 %>% filter (Country.Region == "Ireland")) +
  geom_text_repel(data = Prepared.20 %>%
                    filter(Day2 == max(Day2)),
                  aes(x=Day2+1.75, y=Sum.Cases, 
                      label=Country.Region), size=4) +
  scale_y_log10(labels=scales::comma) +
  ylab('Total number of cases') + xlab('Days since 20th case') +
  ggtitle('Ireland COVID-19 in context') +
  theme_grey(base_size = 22) +
  theme(legend.position="none")

g

ggsave('images/Total_Observed_latest.png', g,
       units='cm', dpi=600, width=21, height=15)

```

# Second plot - Ireland and predictions in context 50 cases

```{r plot2}
Prepared2 <- Prepared  %>%
                    filter(Country.Region %in%
                             c("Ireland", "Italy", "Japan", "Korea, South",
                               "Netherlands", "Spain", "United Kingdom",
                               "Singapore", "New Zealand", "Australia"))


Prepared2.50 <- Prepared2 %>%
  group_by(Country.Region) %>%
  filter(Sum.Cases > 50) %>%
  mutate(Day2 = seq_along(date))

Results.50 <- Results %>%
  filter(Sum.Cases > 50 | is.na(Sum.Cases)) %>%
  mutate(Day2 = seq_along(date))


g <- ggplot(data=Prepared2.50,
            aes(x=Day2, y=Sum.Cases, colour=Country.Region)) +
  geom_path() +
  geom_point(data=Prepared2.50 %>% filter(Country.Region == 'Ireland'),
             aes(x=Day2,y=Sum.Cases),
                 colour='red') +
  geom_path(data=Results.50,
            aes(y=Prediction.Gom, x=Day2),
                 colour='blue') +
  geom_ribbon(data=Results.50,
            aes(ymin=Lower, ymax=Upper),
            colour='pink', fill='green', alpha=0.5) +
  ylab('Total number of cases') + xlab('Days since fiftieth case') +
  ggtitle('Ireland COVID-19 in context',
          subtitle = 'Data and predictions') +
  geom_text_repel(data = Prepared2.50 %>%
                    filter(Day2 == max(Day2)),
                  aes(x=Day2, y=0.8*Sum.Cases, 
                      label=Country.Region),size=4) +
  geom_label_repel(data=Results.50 %>%
                     filter(Day == max(Day)),
             aes(x=tail(Results.50$Day2,1)+1, y=0.6*tail(Results.50$Prediction.Gom,1),
                       label='Predicted'),size=3,colour='pink') +
  scale_y_log10(labels=scales::comma) +
  scale_x_continuous() +
  theme_grey(base_size = 22) +
  theme(legend.position="none")

g

ggsave('images/Total_predictions_latest.png', g,
       units='cm', dpi=600, width=21, height=15)
```

## Graphs of latest daily predictions

```{r}
Prediction <- Results %>%
  mutate(Daily.Cases.Gom = Prediction.Gom - lag(Prediction.Gom)) %>% 
  mutate(Daily.Cases.Lower = Lower.Gom - lag(Lower.Gom)) %>% 
  mutate(Daily.Cases.Upper = Upper.Gom - lag(Upper.Gom)) %>%
  mutate(Observed = Total.Cases) %>%
  mutate(Predicted = Daily.Cases.Gom) %>%
  mutate(Date = (Day -1) + ymd('2020-03-03')) 

g <- ggplot(data = Prediction)  +
  geom_point(aes(x = Date, y = Observed),colour='red') +
  geom_line(aes(x = Date, y = Daily.Cases.Gom),colour='blue',alpha=0.4) +
  geom_ribbon(aes(x=Date, ymin=Daily.Cases.Lower, ymax=Daily.Cases.Upper),
            colour='blue', fill='lightblue', alpha=0.3) +
  geom_point(aes(x = Date, y = Predicted),colour='blue',alpha=0.4) +
  ylab('Daily reported cases') +
  ggtitle('Reported daily cases of COVID-19 infection',
          subtitle = 'Observed and predicted (Gompertz model)') +
  theme_grey(base_size = 22)  +
  scale_x_date(date_breaks = "1 week", date_labels = "%d-%m")

g

ggsave('images/Daily_predictions_latest_Gom.png',g,units='cm',dpi=600,width=21,height=15)
```


```{r}
Prediction <- Results %>%
  mutate(Daily.Cases.LL = Prediction.LL - lag(Prediction.LL)) %>% 
  mutate(Daily.Cases.Lower = Lower.LL - lag(Lower.LL)) %>% 
  mutate(Daily.Cases.Upper = Upper.LL - lag(Upper.LL)) %>%
  mutate(Observed = Total.Cases) %>%
  mutate(Predicted = Daily.Cases.LL) %>%
  mutate(Date = (Day -1) + ymd('2020-03-03')) 

g <- ggplot(data = Prediction)  +
  geom_point(aes(x = Date, y = Observed),colour='red') +
  geom_line(aes(x = Date, y = Daily.Cases.LL),colour='blue',alpha=0.4) +
  geom_ribbon(aes(x=Date, ymin=Daily.Cases.Lower, ymax=Daily.Cases.Upper),
            colour='blue', fill='lightblue', alpha=0.3) +
  geom_point(aes(x = Date, y = Predicted),colour='blue',alpha=0.4) +
  ylab('Daily reported cases') +
  ggtitle('Reported daily cases of COVID-19 infection',
          subtitle = 'Observed and predicted LL5 model') +
  theme_grey(base_size = 22)  +
  scale_x_date(date_breaks = "2 week", date_labels = "%d-%m")
g

ggsave('images/Daily_predictions_latest_LL.png',g,units='cm',dpi=600,width=21,height=15)
```

