---
title: "Load data"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: 
    fig_caption: yes
    toc: yes
editor_options:
  chunk_output_type: console
---

# Purpose
This file is run once a day to download the latest coronavirus data. The Irish data, manually collected each evening from the DoH website (https://www.gov.ie/en/news/7e0924-latest-updates-on-covid-19-coronavirus/) are downloaded separately, and used to correct the JHU data. Need to see if there are better resources than JHU.

```{r setup, include=FALSE}
rm(list=ls())
library(devtools)
library(lubridate)
library(jsonlite)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Update
Load the latest accumulated international coronavirus data from GitHub

```{r load}
install_github("RamiKrispin/coronavirus")

```

Check to see if it worked.

```{r test}
library(coronavirus)
data("coronavirus")
glimpse(coronavirus)

Ireland_JHU <- coronavirus %>%
  filter(country == 'Ireland', type == 'confirmed') 

```

Do a quick table or two

```{r some tables}
summary_df <- coronavirus %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases)

coronavirus %>% 
  filter(date == max(date)) %>%
  select(country = country, type, cases) %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type,
              values_from = total_cases) %>%
  arrange(-confirmed)
```

## Correct the Irish data

The Irish data contain some errors in the JHU dataset, and are less up-to-date. We download the data from the OG website

# CovidStatisticsProfileHPSCIrelandOpenData
https://covid-19.geohive.ie/datasets/d8eb52d56273413b84b0187a4e9117be

Two datasets mashed together. One up to date, the other two days behind

```{r Ireland}
Irish_Open_data <- read_csv('https://opendata.arcgis.com/datasets/d8eb52d56273413b84b0187a4e9117be_0.csv')
```

## Irish current data

```{r Irish current data}
Irish_Current <- Irish_Open_data %>% dplyr::select(X:TotalCovidDeaths,FID) %>%
  mutate(Date = as_date(Date))

# fix Irish current if needed
#Add <- tail(Irish_Current,n=1)
#Add
#Add$Date <- Add$Date+1
#Add$ConfirmedCovidCases = 1718
#Add$TotalConfirmedCovidCases = Add$TotalConfirmedCovidCases + 1718
#Add
#Irish_Current = rbind(Irish_Current,Add)
saveRDS(Irish_Current, file='data/Irish_Current.rds')

```

## Irish Profile data

```{r}
Irish_Profile <- Irish_Open_data %>% dplyr::select(StatisticsProfileDate:FID) %>%
  mutate(province = NA_character_, country='Ireland') %>%
  rename(date = StatisticsProfileDate) %>%
  mutate(Date = as_date(date)) %>%
  select(-date) %>%
  select(Date,CovidCasesConfirmed:country)


Irish_Profile <- Irish_Profile %>%
  mutate(Admissions = HospitalisedCovidCases -
           lag(HospitalisedCovidCases, default = 0)) %>%
  mutate(Admissions = ifelse(Admissions < 0, 0, Admissions))

Irish_Profile <- Irish_Profile %>%
  mutate(ICU = RequiringICUCovidCases -
           lag(RequiringICUCovidCases, default = 0)) %>%
  mutate(ICU = ifelse(ICU < 0, 0, ICU))

saveRDS(Irish_Profile, file='data/Irish_Profile.rds')
```

# County level data

```{r}
CountyCSV <- read_csv('http://opendata-geohive.hub.arcgis.com/datasets/d9be85b30d7748b5b7c09450b8aede63_0.csv?outSR={%22latestWkid%22:3857,%22wkid%22:102100}')
rm(CountyCSV)
CountyCSV <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/d9be85b30d7748b5b7c09450b8aede63_0.csv')

saveRDS(CountyCSV, file='data/CountyCSV.rds')
```

# Covid19AcuteHospitalHistoricSummaryOpenData 

From :- https://data.gov.ie/dataset/covid19acutehospitalhistoricsummaryopendata1/resource/83999570-3323-4ff4-b9a3-f5eaaa3c33b7

```{r}
Hospital_SummaryCSV <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/fe9bb23592ec4142a4f4c2c9bd32f749_0.csv?outSR=%7B%22latestWkid%22%3A4326%2C%22wkid%22%3A4326%7D')

saveRDS(Hospital_SummaryCSV, file='data/Hospital_SummaryCSV.rds')
```

## LaboratoryLocalTimeSeriesHistoricView

From :- https://data.gov.ie/dataset/laboratorylocaltimeserieshistoricview1/resource/0d9e8b56-1aac-464a-bb39-38c6e5697bb7

```{r}
Lab_RecordsCSV <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/f6d6332820ca466999dbd852f6ad4d5a_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D')

# This file is cumulative data only
# Make daily data
Lab_Records <- Lab_RecordsCSV %>%
  mutate(Date = as_date(Date_HPSC)) %>%
  mutate(TestsDone = TotalLabs -
           lag(TotalLabs, default = 0)) %>%
  mutate(TestsPos = Positive -
           lag(Positive, default = 0)) %>%
  mutate(PosRate = 100*round(TestsPos/TestsDone,3))

saveRDS(Lab_Records, file='data/Lab_Records.rds')

```

## Fix JHU data
```{r}

# This file has no rows for days when there are no cases.
  start_date = min(coronavirus$date)
  end_date = max(as_date(Irish_Current$Date))
  n_days <- lubridate::interval(start_date,end_date)/days(1)
Datespan <- as_tibble(start_date + days(0:n_days)) %>%
  rename(date = value) # I row per date in the two datasources

Ireland_corrected <- Irish_Current %>%
  mutate(province = '', country='Ireland') %>%
  rename(date = Date, province =  province, country = country,
         lat = X, long = Y, confirmed = ConfirmedCovidCases,
         Total.Cases = TotalConfirmedCovidCases,
         death = ConfirmedCovidDeaths,
         total.deaths = TotalCovidDeaths) %>%
  mutate(date = as_date(date)) %>%
  dplyr::select(-starts_with('Total.')) %>%
  dplyr::select(date, province, country, lat, long,
         confirmed, death) %>%
  right_join(Datespan, by = "date") %>% # Fill in the dates where no cases occurred in Ireland
  mutate(country = 'Ireland',
         lat = max(Ireland_JHU$lat),
         long= max (Ireland_JHU$long)) %>%
  arrange(date) %>% # Fill in country, lat, long etc..
  pivot_longer(-c(date, province, country, lat, long),
               values_to = 'cases',
               names_to = 'type')

table(Ireland_corrected$province, Ireland_corrected$country, useNA = 'ifany')


# Replace NA with 0 in cases and province (comes from merge with Datespan)
Ireland_corrected$cases[is.na(Ireland_corrected$cases)] <- 0
Ireland_corrected$province[is.na(Ireland_corrected$province)] <- ''


Ireland_corrected %>%
  group_by(type) %>%
  summarise(N = sum(cases))
```

# Fix JHU data

```{r Fix coronavirus data}
# Add back Irish data
JHU_C <- coronavirus %>%
  filter(country != 'Ireland') %>%
  rbind(Ireland_corrected)

Ireland_fixed <- JHU_C %>%
  filter(country == 'Ireland', type == 'confirmed')

saveRDS(JHU_C, file='data/JHU_C.rds')

```

