---
title: "Load data"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: 
    fig_caption: yes
    toc: yes
editor_options:
  chunk_output_type: console
---

# Purpose

This file is run once a day to download the latest coronavirus data. The Irish data, manually collected each evening from the DoH website (https://www.gov.ie/en/news/7e0924-latest-updates-on-covid-19-coronavirus/) are downloaded separately, and used to correct the JHU data.


```{r setup, include=FALSE}
rm(list=ls())
library(devtools)
library(lubridate)
library(jsonlite)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Update
Load the latest accumulated international coronavirus data from GitHub

```{r load}
install_github("RamiKrispin/coronavirus")

```

Check to see if it worked.

```{r test}
library(coronavirus)
data("coronavirus")
glimpse(coronavirus)

Ireland_JHU <- coronavirus %>%
  filter(country == 'Ireland', type == 'confirmed') 

```

Do a quick table or two

```{r some tables}
summary_df <- coronavirus %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases)

coronavirus %>% 
  filter(date == max(date)) %>%
  select(country = country, type, cases) %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type,
              values_from = total_cases) %>%
  arrange(-confirmed)
```

## Correct the Irish data

The Irish data contain some errors in the JHU dataset, and are less up-to-date. We download the data from the OG website

# CovidStatisticsProfileHPSCIrelandOpenData
https://covid-19.geohive.ie/datasets/d8eb52d56273413b84b0187a4e9117be

Two datasets mashed together. One up to date, the other two days behind

```{r Ireland}
Irish_Open_data <- read_csv('https://opendata.arcgis.com/datasets/d8eb52d56273413b84b0187a4e9117be_0.csv') %>%
  mutate(DateD = as_date(Date)) %>% # Up to date
  mutate(StatisticsProfileDateD = as_date(StatisticsProfileDate)) # Two days behind
  
glimpse(Irish_Open_data)
```

## Fix deaths

```{r FixDeaths function}

FixDeaths <- function(DateStart,DateEnd,Dead,DF) {
  cat('DateStart - ', DateStart,'\n',
      'DateEnd - ', DateEnd,'\n',
      'Dead - ', Dead,'\n')
  DateRange <- as_date((as_date(DateStart) : as_date(DateEnd)))
  lamda = Dead/length(DateRange)
    SUM = 0
  while(SUM != Dead){
    Deaths <- rpois(length(DateRange),lamda)
    SUM = sum(Deaths)
    #cat(SUM,'\n')
  }

    FIX <- tibble(DateD = DateRange,
                  ConfirmedCovidDeaths = Deaths)
    
    DF <- DF %>%
      rows_update(FIX, by=c("DateD")) %>%
      arrange(DateD)

    return(DF)
}

```

There were zero reported deaths each day from 2021-05-15 to 2012-06-16. On 2021-06-17 38 deaths were recorded. Thereafter deaths were reported each week. We assume that these 38 deaths came randomly over those days.

```{r fix deaths - first batch ?Ransomware hole '2021-05-15' to '2021-06-17'}
TEST <- FixDeaths(DateStart = '2021-05-15', DateEnd = '2021-06-17',
                  Dead = 38, DF = Irish_Open_data)
sum(TEST$ConfirmedCovidDeaths,na.rm=TRUE)
```

On 2021-06-23 there were 10 deaths reported with missing values for the previous 5 days (6 day interval), and on 2021-07-01 there were 2 deaths reported with 9 the previous day, and six missing values preceding those. We take those 2 as being reported that day, and the six deaths reported on 2021-06-30 as relating to that day and the previous six days back to 2021-06-24. Thereafter (so far) deaths are reported weekly.

```{r assign deaths second batch '2021-06-18' to '2021-06-23'}
TEST <- FixDeaths(DateStart = '2021-06-18', DateEnd = '2021-06-23',
                  Dead = 10, DF = TEST)
sum(TEST$ConfirmedCovidDeaths,na.rm=TRUE)
```

```{r assign deaths third batch '2021-06-24' to '2021-06-30'}
TEST <- FixDeaths(DateStart = '2021-06-24', DateEnd = '2021-06-30',
                  Dead = 9, DF = TEST)
sum(TEST$ConfirmedCovidDeaths,na.rm=TRUE)

```

```{r assign deaths fourth batch '2021-07-02' to '2021-07-07'}
TEST <- FixDeaths(DateStart = '2021-07-02', DateEnd = '2021-07-07',
                  Dead = 6, DF = TEST)
sum(TEST$ConfirmedCovidDeaths,na.rm=TRUE)

```

Thereafter (so far) deaths are reported weekly.

```{r}
TEST_subset <- TEST %>%
  select(DateD,ConfirmedCovidDeaths) %>%
  filter(DateD > as_date('2021-07-07')) %>%
  filter(!is.na(ConfirmedCovidDeaths)) %>%
  mutate(DIFF = DateD - as_date('2021-07-07')) %>%
  mutate(DateEnd = DateD) %>%
  mutate(DateStart = DateEnd - 6)
glimpse(TEST_subset)

for (i in seq(1:length(TEST_subset$DateD))){
     cat(i,'\n')
TEST <- FixDeaths(DateStart = TEST_subset$DateStart[i],
            DateEnd = TEST_subset$DateEnd[i],
            Dead = TEST_subset$ConfirmedCovidDeaths[i],
            DF=TEST)
}

warnings()
sum(TEST$ConfirmedCovidDeaths,na.rm=TRUE)

Irish_Open_data <- TEST

```


## Irish current data

```{r Irish current data}
Irish_Current <- Irish_Open_data %>% dplyr::select(X:TotalCovidDeaths,FID) %>%
  mutate(Date = as_date(Date))

# fix Irish current if needed
#Add <- tail(Irish_Current,n=1)
#Add
#Add$Date <- Add$Date+1
#Add$ConfirmedCovidCases = 6888
#Add$TotalConfirmedCovidCases = #Add$TotalConfirmedCovidCases + 6888
#Add
#Irish_Current = rbind(Irish_Current,Add)
saveRDS(Irish_Current, file='data/Irish_Current.rds')

```

## Irish Profile data

```{r}
Irish_Profile <- Irish_Open_data %>% dplyr::select(StatisticsProfileDate:Aged85up) %>%
  mutate(province = NA_character_, country='Ireland') %>%
  rename(date = StatisticsProfileDate) %>%
  mutate(Date = as_date(date)) %>%
  select(-date) %>%
  select(Date,CovidCasesConfirmed:Aged85up)

Irish_Profile <- Irish_Profile %>%
  mutate(Admissions = HospitalisedCovidCases -
           lag(HospitalisedCovidCases, default = 0)) %>%
  mutate(Admissions = ifelse(Admissions < 0, 0, Admissions))

Irish_Profile <- Irish_Profile %>%
  mutate(ICU = RequiringICUCovidCases -
           lag(RequiringICUCovidCases, default = 0)) %>%
  mutate(ICU = ifelse(ICU < 0, 0, ICU))
```


Save the Irish_Profile

```{r}
saveRDS(Irish_Profile, file='data/Irish_Profile.rds')
```

# County level data

```{r}
CountyCSV <- read_csv('http://opendata-geohive.hub.arcgis.com/datasets/d9be85b30d7748b5b7c09450b8aede63_0.csv?outSR={%22latestWkid%22:3857,%22wkid%22:102100}')
CountyCSV2 <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/d9be85b30d7748b5b7c09450b8aede63_0.csv')
rm(CountyCSV2)

CountyCSV <- CountyCSV %>%
  mutate(Date = as_date(TimeStamp)) %>%
  filter(Date > as_date('2020-03-20')) # No data till 2020-03-21


saveRDS(CountyCSV, file='data/CountyCSV.rds')
```

# Covid19AcuteHospitalHistoricSummaryOpenData 

From :- https://data.gov.ie/dataset/covid19acutehospitalhistoricsummaryopendata1/resource/83999570-3323-4ff4-b9a3-f5eaaa3c33b7

```{r}
Hospital_SummaryCSV <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/fe9bb23592ec4142a4f4c2c9bd32f749_0.csv?outSR=%7B%22latestWkid%22%3A4326%2C%22wkid%22%3A4326%7D')

saveRDS(Hospital_SummaryCSV, file='data/Hospital_SummaryCSV.rds')
```

## LaboratoryLocalTimeSeriesHistoricView

From :- https://data.gov.ie/dataset/laboratorylocaltimeserieshistoricview1/resource/0d9e8b56-1aac-464a-bb39-38c6e5697bb7

```{r}
Lab_RecordsCSV <- read_csv('https://opendata-geohive.hub.arcgis.com/datasets/f6d6332820ca466999dbd852f6ad4d5a_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D')

# This file is cumulative data only
# Make daily data
Lab_Records <- Lab_RecordsCSV %>%
  mutate(Date = as_date(Date_HPSC)) %>%
  mutate(TestsDone = TotalLabs -
           lag(TotalLabs, default = 0)) %>%
  mutate(TestsPos = Positive -
           lag(Positive, default = 0)) %>%
  mutate(PosRate = 100*round(TestsPos/TestsDone,3))

saveRDS(Lab_Records, file='data/Lab_Records.rds')

```

## Fix JHU data
```{r}

# This file has no rows for days when there are no cases.
  start_date = min(coronavirus$date)
  end_date = max(as_date(Irish_Current$Date))
  n_days <- lubridate::interval(start_date,end_date)/days(1)
  
Datespan <- as_tibble(start_date + days(0:n_days)) %>%
  rename(date = value) # I row per date in the two datasources

Ireland_corrected <- Irish_Current %>%
  mutate(province = '', country='Ireland') %>%
  rename(date = Date, province =  province, country = country,
         lat = X, long = Y, confirmed = ConfirmedCovidCases,
         Total.Cases = TotalConfirmedCovidCases,
         death = ConfirmedCovidDeaths,
         total.deaths = TotalCovidDeaths) %>%
  mutate(date = as_date(date)) %>%
  dplyr::select(-starts_with('Total.')) %>%
  dplyr::select(date, province, country, lat, long,
         confirmed, death) %>%
  right_join(Datespan, by = "date") %>% # Fill in the dates where no cases occurred in Ireland
  mutate(country = 'Ireland',
         lat = max(Ireland_JHU$lat),
         long= max (Ireland_JHU$long)) %>%
  arrange(date) %>% # Fill in country, lat, long etc..
  pivot_longer(-c(date, province, country, lat, long),
               values_to = 'cases',
               names_to = 'type') %>%
mutate(uid = NA*cases, iso2 = NA*cases, iso3 = NA*cases, code3 = NA*cases, combined_key = NA*cases, population = NA*cases, continent_name = NA*cases, continent_code = NA*cases)
table(Ireland_corrected$province, Ireland_corrected$country, useNA = 'ifany')


# Replace NA with 0 in cases and province (comes from merge with Datespan)
Ireland_corrected$cases[is.na(Ireland_corrected$cases)] <- 0
Ireland_corrected$province[is.na(Ireland_corrected$province)] <- ''


Ireland_corrected %>%
  group_by(type) %>%
  summarise(N = sum(cases))
```

# Fix JHU data

```{r Fix coronavirus data}
# Add back Irish data
JHU_C <- coronavirus %>%
  filter(country != 'Ireland') %>%
  rbind(Ireland_corrected)

Ireland_fixed <- JHU_C %>%
  filter(country == 'Ireland', type == 'confirmed')

saveRDS(JHU_C, file='data/JHU_C.rds')

```


## OWID data

```{r,eval=FALSE}
OWID <- read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv", guess_max=100000)
  str(OWID)
  
Population <- OWID %>% select(iso_code,location,population) %>% distinct()
  saveRDS(OWID, file='data/OWID.rds')
  saveRDS(Population, file='data/Population.rds')
```


# Manual update of Irish Data

```{r,eval=FALSE}
X <- Irish_Current$X[1]
Y <- Irish_Current$Y[1]

start_Date <- max(Irish_Current$Date) + 1 
end_Date <- Sys.Date()
Span <- days(end_Date - start_Date)@day
Date <- seq(start_Date, end_Date, 1)
ConfirmedCovidCases <- c(447,355,360,358,503,469,
                         524,381,438,345,365,448,
                         436,467,464,374,378,337,
                         407,465,529,416,313,377)


TotalConfirmedCovidCases <- ConfirmedCovidDeaths <- TotalCovidDeaths <- FID <- 0  
Rows = tibble(X,Y,Date,ConfirmedCovidCases,TotalConfirmedCovidCases,ConfirmedCovidDeaths,TotalCovidDeaths,FID)

Irish_Current <- Irish_Current %>% rbind(Rows)

saveRDS(Irish_Current, file='data/Irish_Current.rds')
```
