---
title: "County data"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#Purpose
To prepare graphs of coronavirus incidence at county level

```{r setup, include=FALSE}
rm(list=ls())
library(sf)
library(RcppRoll)
library(scales)
library(lubridate)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```


```{r load data file}
CountyCSV <- readRDS('data/CountyCSV.rds')

```

## Basic dates

```{r}
CountyCSV <- CountyCSV %>%
  mutate(Date = as_date(TimeStamp)) %>%
  filter(Date > as_date('2020-03-20')) # No data till 2020-03-21

start_date = min(as_date(CountyCSV$Date))
end_date = max(as_date(CountyCSV$Date))
  n_days <- interval(start_date,end_date)/days(1)
Datespan <- as_tibble(start_date + days(0:n_days)) %>%
  rename(Date = value) # I row per date in the two datasources
```

County level data has cumulative cases, not cases per day, which is what we need, so we get the difference of the cases on each day, from the previous day, within each county.

```{r}
#Lags and cases
CountyCSV <- CountyCSV %>%
  group_by(ORIGID) %>%
  mutate(Cases = ConfirmedCovidCases -
           lag(ConfirmedCovidCases, default = 0)) %>%
  mutate(Cases = ifelse(Cases <0,0,Cases))

#Means and Medians
CountyCSV <- CountyCSV %>%
  group_by(ORIGID) %>%
  mutate(CasesSum7 = roll_sum(Cases,n=7,fill=0,align='right')) %>%
  mutate(CasesSum14 = roll_sum(Cases,n=14,fill=0,align='right')) %>%
  mutate(CasesMean7 = roll_mean(Cases,n=7,fill=0,align='right')) %>%
  mutate(CasesMean14 = roll_mean(Cases,n=14,fill=0,align='right')) %>%
  mutate(CasesMedian7 = roll_median(Cases,n=7,fill=0,align='right')) %>%
  mutate(CasesMedian14 = roll_median(Cases,n=14,fill=0,align='right'))

#Rates
CountyCSV <- CountyCSV %>%
  group_by(ORIGID) %>%
  mutate(Rate = 100000*Cases/PopulationCensus16) %>%
  mutate(MeanRate7 = 100000*CasesMean7/PopulationCensus16) %>%
  mutate(MeanRate14 = 100000*CasesMean14/PopulationCensus16) %>%
  mutate(MedianRate7 = 100000*CasesMean7/PopulationCensus16) %>%
  mutate(MedianRate14 = 100000*CasesMedian14/PopulationCensus16)

write_csv(CountyCSV,'CountyCSV.csv')

Kil <- CountyCSV %>%
  filter(ORIGID == 9) # Kildare

Dub <- CountyCSV %>%
  filter(ORIGID == 6) # Dublin
```

## Load the boundaries

```{r}
Counties <- st_read('County boundaries/Census2011_Admin_Counties_generalised20m.shp')
Counties
Counties$COUNTYNAME[20] <- 'Dún Laoghaire-Rathdown' # Fix typo

ggplot(data=Counties) + geom_sf() + ggtitle('Republic of Ireland')  + geom_sf_text(aes(label = COUNTYNAME))
```

## Load the geography matching file

```{r}
Match_Geographies <- openxlsx::read.xlsx(
  'County boundaries/Merge_Geographies.xlsx')

CountyCSV <- CountyCSV %>%
  inner_join(Match_Geographies %>% select(Data,Abbreviation) %>% distinct(),
             by=c('CountyName' = 'Data' ))
```


## Heatmap
county by date
Order by date and total rate
Display count

```{r, fig.width=8, fig.height=20}
#CountyOrder <- CountyCSV %>% group_by(Abbreviation) %>% summarise(MAX = max(ConfirmedCovidCases)) %>% arrange(desc(MAX))
CountyOrder <- CountyCSV %>%
  group_by(Abbreviation) %>%
  summarise(MAX = max(MeanRate14)) %>%
  arrange(desc(MAX))

ORDER <- as.character(CountyOrder$Abbreviation)

CountyCSV <- CountyCSV %>%
  mutate(AbbreviationF = as_factor(Abbreviation)) %>%
  mutate(AbbreviationF =  fct_relevel(AbbreviationF,ORDER))

ggplot(data=CountyCSV,
       aes(x=AbbreviationF,y=Date)) +
  geom_tile(aes(colour = MeanRate14,
                fill = MeanRate14),height=1,width=1) +
  scale_x_discrete(position = "top") +
  scale_colour_distiller(palette='OrRd', direction=1,
                         na.value='white') +
  scale_fill_distiller(palette='OrRd', direction=1,
                         na.value='white') +
#  scale_color_manual(values=HeatMapColours,
#    aesthetics = c("colour", "fill")) +
  scale_y_date(breaks='1 month',
  minor_breaks =  "1 week",
  labels=date_format("%b")) +
  theme(axis.text.x.top = element_text(angle = 270, face='bold',
                                       vjust = 0.5, hjust=1,
                                       size=20)) +
  labs(title =  '7 day case rate by county',
       x = 'County',  y = 'Date',
       colour='Rate', fill='Rate') +
  theme_light()

```
