---
title: "Time trends"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: 
    fig_caption: yes
    toc: yes
editor_options:
  chunk_output_type: console
---

#Purpose
To prepare graphs of coronavirus incidence over time.

```{r setup, include=FALSE}
rm(list=ls())

library(RcppRoll)
library(lubridate)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

# Load the data
Latest HSE/HPSC data

```{r}
Irish_Current <- readRDS('data/Irish_Current.rds')
  start_date = min(Irish_Current$Date)
  end_date = max(Irish_Current$Date) + days(5)

Lab_Records <-  readRDS('data/Lab_Records.rds')
min(Lab_Records$Date_HPSC)
max(Lab_Records$Date_HPSC)
```

# Cases over time

## Since the start

```{r}
ggCases <- ggplot(Irish_Current,aes(x=date(Date),y=ConfirmedCovidCases)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Confirmed COVID19 cases over time') +
  theme_light()

ggCases

ggsave('images/COVID19Caseslogscale.png', ggCases,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

## Last 60 days

```{r}
ggCases <- ggplot(Irish_Current %>% filter(Date > end_date-60),
                  aes(x=date(Date),y=ConfirmedCovidCases)) +
  geom_path(alpha=0.3) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.5,colour='blue') +
  scale_x_date(limits=c(end_date-60, end_date),
               date_breaks = "1 week",
               date_labels = "%d-%b",
               date_minor_breaks = "1 day") +
#  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Confirmed COVID19 cases over time') +
  theme_light()

ggCases

ggsave('images/COVID19CasesShortlinear.png', ggCases,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

# Deaths

## Since the start

```{r}
ggDeaths <- ggplot(Irish_Current,aes(x=date(Date),y=ConfirmedCovidDeaths)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Confirmed COVID19 deaths over time') +
  theme_light()

ggDeaths

ggsave('images/COVID19Deathslogscale.png', ggDeaths,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

## Last 60 days

```{r}
ggDeaths <- ggplot(Irish_Current,aes(x=date(Date),y=ConfirmedCovidDeaths)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(end_date-60, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Confirmed COVID19 deaths over time') +
  theme_light()

ggDeaths

ggsave('images/COVID19Deathslogscale.png', ggDeaths,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

# Load the profile data

```{r}
Irish_Profile <- readRDS('data/Irish_Profile.rds') %>%
  filter(!is.na(CovidCasesConfirmed))

start_date = min(Irish_Profile$Date)
end_date = max(Irish_Profile$Date) + days(5)

```

# Admissions

## Since the start

```{r}
ggAdmissions <- ggplot(Irish_Profile,aes(x=date(Date),y=Admissions)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date-22),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Hospitalised COVID19 Cases over time') +
  theme_light()

ggAdmissions

ggsave('images/COVID19Admissionslogscale.png', ggAdmissions,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

## Last 60 days

```{r}
ggAdmissions <- ggplot(Irish_Profile,aes(x=date(Date),y=Admissions)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(end_date-60, end_date-22),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_log10() +
  xlab('Date') + ylab('Cases (log scale)') +
  ggtitle('Hospitalised COVID19 Cases over time') +
  theme_light()

ggAdmissions

ggsave('images/COVID19Admissionslogscale.png', ggAdmissions,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

# ICU Admissions

## Since the start

```{r}
ggICU <- ggplot(Irish_Profile,aes(x=date(Date),y=ICU)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
#  scale_y_log10() +
  xlab('Date') + ylab('Cases') +
  ggtitle('Hospitalised COVID19 Cases in ICU over time') +
  theme_light()

ggICU

ggsave('images/COVID19ICUlogscale.png', ggICU,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

# Laboratory data

```{r}
Lab_Records <- readRDS(file='data/Lab_Records.rds')

```

## Number of tests done

### since start

```{r}
ggTests <- ggplot(Lab_Records,aes(x=date(Date),y=TestsDone)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
#  scale_y_log10() +
  xlab('Date') + ylab('Tests') +
  ggtitle('Laboratory tests done') +
  theme_light()

ggTests

ggsave('images/LabTestsDone.png', ggTests,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

### Last 60 days

```{r}
ggTests <- ggplot(Lab_Records,aes(x=date(Date),y=TestsDone)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(end_date-60, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
#  scale_y_log10() +
  xlab('Date') + ylab('Tests') +
  ggtitle('Laboratory tests done') +
  theme_light()

ggTests

ggsave('images/LabTestsDone.png', ggTests,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

## Tests positive

### Since start

```{r}
ggTestsPos <- ggplot(Lab_Records,aes(x=date(Date),y=TestsPos)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
#  scale_y_log10() +
  xlab('Date') + ylab('Tests') +
  ggtitle('Positive tests number') +
  theme_light()

ggTestsPos

ggsave('images/LabTestsPos.png', ggTestsPos,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

### Last 60 days

```{r}
ggTestsPos <- ggplot(Lab_Records,aes(x=date(Date),y=TestsPos)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(end_date -60, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
#  scale_y_log10() +
  xlab('Date') + ylab('Tests') +
  ggtitle('Positive tests number') +
  theme_light()

ggTestsPos

ggsave('images/LabTestsPos.png', ggTestsPos,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```


## Percentage tests positive

### Since start

```{r}
ggTestsPosPct <- ggplot(Lab_Records,aes(x=date(Date),y=PosRate)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(start_date, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  xlab('Date') + ylab('Percentage tests positive') +
  ggtitle('Positive tests %') +
  theme_light()

ggTestsPosPct

ggsave('images/LabTestsPosPct.png', ggTestsPosPct,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

### Last 60 days

```{r}
ggTestsPosPct <- ggplot(Lab_Records,aes(x=date(Date),y=PosRate)) +
  geom_path(alpha=0.1) +
  geom_smooth(alpha=0.1,span=0.2,se=FALSE,colour='pink') +
  geom_point(alpha=0.3,colour='blue') +
  scale_x_date(limits=c(end_date - 60, end_date),
               date_breaks = "1 month",
               date_labels = "%b",
               date_minor_breaks = "1 week") +
  scale_y_continuous(limits=c(0,15.0)) +
  xlab('Date') + ylab('Percentage tests positive') +
  ggtitle('Positive tests %') +
  theme_light()

ggTestsPosPct

ggsave('images/LabTestsPosPct.png', ggTestsPosPct,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```



