---
title: "Vaccine Estimates"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: 
    fig_caption: yes
    toc: yes
editor_options:
  chunk_output_type: console
---

#Purpose

To examine a few simple deterministic scenarios for vaccine roll out for the new COVID19 vaccines.

```{r setup, include=FALSE}
rm(list=ls())

library(lubridate) # Sane date handling
library(tidyverse)


knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning=FALSE, results = 'hide')
```

#Parameters

```{r}
POP = 4970000 # Total population
COVER = 0.8   # Required coverage
DOSES = POP*COVER*2 # Number of doses required

```

# Starting assumptions

We have `r POP` people in Ireland.
We need to vaccinate `r POP*COVER` of these for 80% coverage.
This requires `r DOSES` doses of vaccine.


# Effect of different fixed vaccination rates

```{r}
Start <- as_date('2021-01-04') #Monday of Week 1
Weeks <- seq(0,265,1)
Dates <- Start + 7 * Weeks


VaccRates <- c(0,seq(50000,200000, 25000)) #7 + 1
VaccRatesNames <- fct_inorder(c('0k','50k', '75k', '100k', '125k', '150k', '175k', '200k')) #7 + 1
```

Suppose we start in Week 1 on the `r Start`.

We examine seven weekly vaccination rates (from Week 1). We ignore supply issues, holidays, and any other real life events. We assume we can do from `r VaccRatesNames[2]` to `r VaccRatesNames[8]` each week.

```{r}
VaccineEst <- tibble(Dates,Weeks,
                       VaccRateName = VaccRatesNames[1],
                       VaccRate = VaccRates[1],
                     Delivered = 0, I = 0)

for (i in seq(2,8,1)) {
  print(paste('i = ', i, VaccRatesNames[i], VaccRates[i],'\n'))
  V <- tibble(Dates,Weeks,
                       VaccRateName=VaccRatesNames[i],
                       VaccRate=VaccRates[i]) %>%
  mutate(Delivered = VaccRate*Weeks) %>%
  mutate(I = i)

  VaccineEst <- rbind(VaccineEst,V)
  rm(V)
 }

VaccineEst <- VaccineEst %>%
  filter(I > 0) %>% # Remove the initial blank data frame
  filter(Delivered <= DOSES) # Cut off when we get to full coverage

```

This gives us the following results :-

```{r}
gVaccFig1 <- ggplot(data=VaccineEst,
       aes(x=Dates,y=100*(DOSES - Delivered)/ DOSES, group = VaccRateName, colour=VaccRateName)) +
  geom_path(alpha=0.1) +
  geom_point(alpha=0.3) +
    scale_x_date(limits=c(Dates[1], Dates[52]),
               date_breaks = "3 month",
               date_labels = "%b-%y",
               date_minor_breaks = "1 month") +
  xlab('Month') + ylab('Percentage to be done') +
  ggtitle('2021 Irish Vaccination coverage by time') +
  guides(colour = guide_legend("Vaccinations\nper week")) + 
  theme_light() +
  theme(panel.grid.major.x = 
          element_line(colour = "red", linetype = "dotted"))

gVaccFig1

ggsave('images/gVaccFig1.png', gVaccFig1,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)

```

Unless we can deliver more than 150,000 doses of vaccine each week from now, we do not reach population coverage before Christmas.

# Scaled up vaccination over times

A more credible model is a scaled up vaccination over time. For simplicty we assume a constant rate of increase each week from 20,000 doses a week.

```{r}
Rate_of_Increase = c(1.0,1.1,1.2,1.3,1.4,1.5,1.6,1.7,1.8,2.0)
  Rate_of_IncreaseName = fct_inorder(c('None','10%','20%','30%','40%','50%','60%','70%','80%','100%'))

DoseStart = 20000
  DoseMax = seq(0,700000,100000)
  DoseMaxName <- fct_inorder(c('0k','100k', '200k', '300k', '400k', '500k', '600k', '700k')) #7 + 1

```

We start by lookign only at the effect of an increasing numebr of vaccinations a week, but with a very large maximum number of disease a week - 500,000.

```{r}
DoseMaxSingle=500000

Scale <- tibble(Dates, Weeks,
                Rate = 0, RateName=Rate_of_IncreaseName[1],
                Ramp =  0, Doses = 0,
                Delivered = 0, I = 0)

  for (i in seq(2,10,1)) {
    print(paste('i = ', i, Rate_of_IncreaseName[i]))
    
    S <- tibble(Dates, Weeks) %>%
      mutate(Rate = Rate_of_Increase[i]) %>%
      mutate(RateName = Rate_of_IncreaseName[i]) %>%
      mutate(Ramp = Rate_of_Increase[i] ^ Weeks) %>%
      mutate(Doses = DoseStart * Ramp) %>%
      mutate(Doses = pmin(Doses, DoseMaxSingle)) %>%
      mutate(Delivered = cumsum(Doses)) %>%
      mutate(I = i)
    Scale <- rbind(Scale,S)
  rm(S)
  }

Scale <- Scale %>%
  filter(I > 0) %>% # Lose inital blank dataframe
  mutate(PerCentToCompletion = 100*(DOSES - Delivered)/DOSES) %>% 
  group_by(Rate) %>%
   # Truncate if there are two consecutive values at or over 100% complete
  mutate(SecondZero = ifelse(PerCentToCompletion <= 0 &
                               lag(PerCentToCompletion <= 0),TRUE,FALSE)) %>%
  filter(!SecondZero) %>%
  #Set the smallest to zero
  mutate(PerCentToCompletion  = ifelse(PerCentToCompletion < 0, 0, PerCentToCompletion))

```

This gives us the following picture :-

```{r}
gVaccFig2 <- ggplot(data=Scale,
       aes(x=Dates,y=PerCentToCompletion,
           group = RateName, colour=RateName)) +
  geom_path(alpha=0.1) +
  geom_point(alpha=0.3) +
  scale_x_date(limits=c(Dates[1], Dates[52]),
               date_breaks = "3 month",
               date_labels = "%b-%y",
               date_minor_breaks = "1 month") +
  xlab('Month') + ylab('Percentage to be done') +
  ggtitle('2021 Irish Vaccination coverage by time',
          subtitle = 'by weekly scale up and maximum weekly number of 500k') +
  guides(colour = guide_legend("Weekly increase in\nvaccinations")) + 
  theme_light()+ theme(panel.grid.major.x = element_line(colour = "red", linetype = "dotted"))

gVaccFig2

ggsave('images/gVaccFig2.png', gVaccFig2,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)

```

In every scenario we complete vaccination by Christmas, but only if we increase vaccinaiton rates by an average of 20% a week, until we are doing 500,000 a week.

If our maximum is lower, say 250,000 we get this.

```{r}
DoseMaxSingle=250000

Scale <- tibble(Dates, Weeks,
                Rate = 0, RateName=Rate_of_IncreaseName[1],
                Ramp =  0, Doses = 0,
                Delivered = 0, I = 0)

  for (i in seq(2,10,1)) {
    print(paste('i = ', i, Rate_of_IncreaseName[i]))
    
    S <- tibble(Dates, Weeks) %>%
      mutate(Rate = Rate_of_Increase[i]) %>%
      mutate(RateName = Rate_of_IncreaseName[i]) %>%
      mutate(Ramp = Rate_of_Increase[i] ^ Weeks) %>%
      mutate(Doses = DoseStart * Ramp) %>%
      mutate(Doses = pmin(Doses, DoseMaxSingle)) %>%
      mutate(Delivered = cumsum(Doses)) %>%
      mutate(I = i)
    Scale <- rbind(Scale,S)
  rm(S)
  }

Scale <- Scale %>%
  filter(I > 0) %>% # Lose inital blank dataframe
  mutate(PerCentToCompletion = 100*(DOSES - Delivered)/DOSES) %>% 
  group_by(Rate) %>%
   # Truncate if there are two consecutive values at or over 100% complete
  mutate(SecondZero = ifelse(PerCentToCompletion <= 0 &
                               lag(PerCentToCompletion <= 0),TRUE,FALSE)) %>%
  filter(!SecondZero) %>%
  #Set the smallest to zero
  mutate(PerCentToCompletion  = ifelse(PerCentToCompletion < 0, 0, PerCentToCompletion))

```


```{r}

gVaccFig3 <- ggplot(data=Scale,
       aes(x=Dates,y=PerCentToCompletion,
           group = RateName, colour=RateName)) +
  geom_path(alpha=0.1) +
  geom_point(alpha=0.3) +
  scale_x_date(limits=c(Dates[1], Dates[52]),
               date_breaks = "3 month",
               date_labels = "%b-%y",
               date_minor_breaks = "1 month") +
  xlab('Month') + ylab('Percentage to be done') +
  ggtitle('2021 Irish Vaccination coverage by time',
          subtitle = 'by weekly scale up and maximum weekly number of 250k') +
  guides(colour = guide_legend("Weekly increase in\nvaccinations")) + 
  theme_light()+ theme(panel.grid.major.x = element_line(colour = "red", linetype = "dotted"))

gVaccFig3

ggsave('images/gVaccFig3.png', gVaccFig3,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)

```

Again, this is positive, and for most scenarios we hit 80% coverage in September.

# Joint effects of rising weekly rates and different maximum vaccination rates

This suggests the merit of looking at both.

```{r}

Scale <- tibble(Dates, Weeks,
                DoseMax = DoseMax[1], DoseMaxName = DoseMaxName[1],
                Rate = 0, RateName=Rate_of_IncreaseName[1],
                Ramp =  0, Doses = 0,
                Delivered = 0, I = 0)

for (j in seq(2,10,1)) {
  for (i in seq(2,8,1)) {
    print(paste('j = ', j, Rate_of_Increase[j], 'i = ', i, DoseMax[i]))
    
    S <- tibble(Dates, Weeks) %>%
      mutate(DoseMax = DoseMax[i]) %>%
      mutate(DoseMaxName = DoseMaxName[i]) %>%
      mutate(Rate = Rate_of_Increase[j]) %>%
      mutate(RateName = Rate_of_IncreaseName[j]) %>%
      mutate(Ramp = Rate_of_Increase[j] ^ Weeks) %>%
      mutate(Doses = DoseStart * Ramp) %>%
      mutate(Doses = pmin(Doses, DoseMax[i])) %>%
      mutate(Delivered = cumsum(Doses)) %>%
      mutate(I = i)
    Scale <- rbind(Scale,S)
  rm(S)
  }
}  

Scale <- Scale %>%
  filter(I > 0) %>% # Lose inital blank dataframe
  mutate(PerCentToCompletion = 100*(DOSES - Delivered)/DOSES) %>% 
  group_by(DoseMax,Rate) %>%
   # Truncate if there are two consecutive values at or over 100% complete
  mutate(SecondZero = ifelse(PerCentToCompletion <= 0 &
                               lag(PerCentToCompletion <= 0),TRUE,FALSE)) %>%
  filter(!SecondZero) %>%
  #Set the smallest to zero
  mutate(PerCentToCompletion  = ifelse(PerCentToCompletion < 0, 0, PerCentToCompletion))

```

This gives us the following complex picture.

```{r}
gVaccFig4 <- ggplot(data=Scale,
       aes(x=Dates,y=PerCentToCompletion,
           group = DoseMaxName, colour=DoseMaxName)) +
  geom_path(alpha=0.1) +
  geom_point(alpha=0.3) +
  scale_x_date(limits=c(Dates[1], Dates[52]),
               date_breaks = "3 month",
               date_labels = "%b-%y",
               date_minor_breaks = "1 month") +
  xlab('Month') + ylab('Percentage to be done') +
  ggtitle('Irish Vaccination coverage by time',
          subtitle = 'by weekly scale up and maximum weekly number') +
  guides(colour = guide_legend("Maximum\nVaccinations\nper week")) + 
  theme_light()+ theme(panel.grid.major.x = element_line(colour = "red", linetype = "dotted")) + facet_wrap(~RateName)

gVaccFig4

ggsave('images/gVaccFig4.png', gVaccFig4,
       device = 'png', width = 20, height = 10, units = 'cm', dpi = 600)
```

This complex graph shows two effects. As the maximum possible number of vaccinations rises, the length of time taken to completion falls.

As the rate of increase of vaccinations per week rises (10% a week, 20% a week, up to a 100% a week, which is doubling every week), the length of time taken to completion falls.

 In almost any scenario in which the maximum number of vaccinations a week is over 200k, vaccination is completed before Christmas 2021, which is good news.
 
 In relaity, none of these scenarios is credible. All assume that there no hitches with supply, and in fact that supply is not a limiting factor. This isnt true. The maximum number of vaccinations is determined by supply, and by capacity to vaccinate, both of whihc are rising. The number of vaccinations a week is determined by the same things. We expect over time, that the number of vaccinations that can be delivered will rise, but we also need to understand that some parts of the population can be reached much faster than others.
 
Real vaccination programs always have a maddeningly slow start-up, and a long tail, before adequate vaccination coverage is reached.