---
title: "Coronavirus Time"
author: "Anthony Staines"
date: "13/03/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#Purpose
To prepare models for coronavirus incidence, and short term predictions, in Ireland

```{r setup, include=FALSE}
rm(list=ls())
library(coronavirus) # Coronavirus data
library(propagate) # Errors in non-linear regression models
library(ggrepel) # Nice labels
library(lubridate) # Sane date handling
library(drc)
library(sandwich)
library(lmtest)

library(RcppRoll)
library(tidyverse)

data("coronavirus")

knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

```{r load data file}
load('data/CC.Rdata')

```

## Identify the areas and the first date of a recorded confirmed case, and of the first recorded death

```{r Country}
Areas <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(country) %>%
  summarise(Total.Cases = sum(cases)) %>%
  arrange(desc(Total.Cases), country)

FirstDates <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(country) %>%
  filter(cases > 0) %>%
  filter(date == min(date)) %>%
#  ungroup() %>%
  dplyr::select(country, date) %>%
  distinct()  %>%
  arrange(date, country)
  
DeathDates <- CC %>%
  filter(type == 'death') %>%
  group_by(country) %>%
  filter(cases > 0) %>%
  filter(date == min(date)) %>%
  dplyr::select(country, First.Date = date) %>%
  distinct()  %>%
  arrange(First.Date, country)

```

# Prepare a general data set of confirmed cases

```{r General data set}

Prepared <- CC %>%
  filter(type == 'confirmed') %>%
  group_by(country, date ) %>%
  summarise(Total.Cases = sum(cases)) %>% # Collapse over Province.State
  right_join(FirstDates) %>%
  arrange(date) %>%
  mutate(Day = seq_along(date)) %>% # Set up days from first case
  filter(Day >= 0) %>%
  arrange(country, Day) %>%
  mutate(Sum.Cases = cumsum(Total.Cases)) # Cumulative total of cases is Sum.Cases

```

## Prepare an Ireland specific dataset

```{r Ireland specific dataset}
Ireland.Prepared <- CC %>%
  filter(country == 'Ireland') %>%
  filter(type == 'confirmed') %>%
  group_by(country, date ) %>%
  summarise(Total.Cases = sum(cases)) %>%
  left_join(FirstDates, by = "country") %>%
  mutate(Day  = seq_along(date)) %>%
  filter(Day >= 0) %>%
  arrange(country, Day) %>%
  mutate(Sum.Cases = cumsum(Total.Cases))

```


```{r define prediction frame}
Model.Data <- Prepared %>%
  filter(country == 'Ireland') %>%
  filter(Sum.Cases > 0) %>% # First Day with 1 case
  mutate(Day.Original = Day) %>%
  mutate(Day = 1 + (Day - min(Day)))

pred.frame=tibble(Day = seq(1 : ( max(Model.Data$Day) + 10))) # 10 days ahead.

```

# Predictions of log(accumulated cases)


## LL model

```{r drc LL}
## LL model

drc.LL.4 <- drm(Sum.Cases ~ Day,
       data =Model.Data, 
                   fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
drc.LL.5 <- drm(Sum.Cases ~ Day,
       data =Model.Data, 
                   fct = LL.5(names = c("Slope", "Lower Limit", "Upper Limit", "ED50", "f")))
summary(drc.LL.4)
summary(drc.LL.5)
AIC(drc.LL.4)
AIC(drc.LL.5)

plot(drc.LL.4)
plot(drc.LL.5)

#coeftest(drc.LL.4, vcov = sandwich)
#predict.drc(drc.LL.4,as.data.frame(pred.frame),vcov. = sandwich)
#coeftest(drc.LL.5, vcov = sandwich)
#predict(drc.LL.5,as.data.frame(pred.frame),vcov. = sandwich)
#model.LL.predicted <-
#  predict(drc.LL.5,as.data.frame(pred.frame),
#        se.fit=TRUE, interval='confidence',
#        vcov. = sandwich, od=FALSE, level=0.95) %>%
#  as_tibble() %>%
#  cbind(pred.frame) %>%
#  rename(Prediction.LL = Prediction) %>%
#  mutate(Lower.LL = Prediction.LL - 1.96*SE) %>%
#  mutate(Upper.LL = Prediction.LL + 1.96*SE) %>%
#  select(Day,Prediction.LL,Lower.LL,Upper.LL)#

```

## Gompertz model

```{r drc Gompertz model}
##Gompertz model

drc.Gom <- drm(Sum.Cases ~ Day,
       data = Model.Data, 
                   fct = G.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
summary(drc.Gom)

AIC(drc.Gom)
plot(drc.Gom)
modelFit(drc.Gom)
coeftest(drc.Gom, vcov = sandwich)
model.Gom.predicted <-
  predict(drc.Gom,as.data.frame(pred.frame),
        se.fit=TRUE, interval='confidence',
        vcov. = sandwich, od=FALSE, level=0.95) %>%
  as_tibble() %>%
  cbind(pred.frame) %>%
  rename(Prediction.Gom = Prediction) %>%
  mutate(Lower.Gom = Prediction.Gom - 1.96*SE) %>%
  mutate(Upper.Gom = Prediction.Gom + 1.96*SE) %>%
  select(Day,Prediction.Gom,Lower.Gom,Upper.Gom)

```

### Create a file of results

```{r Pull together results}
Results <- Model.Data %>%
  ungroup() %>%
  filter(country=='Ireland')
```

## Very quick plot of data

```{r plot1}
ggplot(data=Prepared,aes(x=Day,y=log10(Sum.Cases),colour=country)) +
  geom_path() +
  geom_point(data=Model.Data,
             aes(x=Day.Original,y=log10(Sum.Cases)),
                 colour='red') +
  theme(legend.position="none")

```

## Plot 1 Ireland against EU  (mostly) 20 cases

```{r}

Prepared.20 <- Prepared %>%
  group_by(country) %>%
  filter(Sum.Cases > 20) %>%
  mutate(Day2 = seq_along(date))

Prepared.20 <- Prepared.20 %>%
  filter(country %in% c('New Zealand', 'Hungary', 'France', 'Spain', 'US', 'United Kingdom','Ireland', 'Norway', 'Denmark', 'Finland', 'Germany'))

g <- ggplot(data=Prepared.20,
       aes(x=Day2,y=Sum.Cases,colour=country)) +
  geom_path() +
  geom_point(data = Prepared.20 %>% filter (country == "Ireland")) +
  geom_text_repel(data = Prepared.20 %>%
                    filter(Day2 == max(Day2)),
                  aes(x=Day2+1.75, y=Sum.Cases, 
                      label=country), size=4) +
  scale_y_log10(labels=scales::comma) +
  ylab('Total number of cases') + xlab('Days since 20th case') +
  ggtitle('Ireland COVID-19 in context') +
  theme_grey(base_size = 22) +
  theme(legend.position="none")

g

ggsave('images/Total_Observed_latest.png', g,
       units='cm', dpi=600, width=21, height=15)

```

# Second plot - Ireland and predictions in context 50 cases

```{r plot2}
Prepared2 <- Prepared  %>%
                    filter(country %in%
                             c("Ireland", "Italy", "Japan", "Korea, South",
                               "Netherlands", "Spain", "United Kingdom",
                               "Singapore", "New Zealand", "Australia"))


Prepared2.50 <- Prepared2 %>%
  group_by(country) %>%
  filter(Sum.Cases > 50) %>%
  mutate(Day2 = seq_along(date))

Results.50 <- Results %>%
  filter(Sum.Cases > 50 | is.na(Sum.Cases)) %>%
  mutate(Day2 = seq_along(date))


g <- ggplot(data=Prepared2.50,
            aes(x=Day2, y=Sum.Cases, colour=country)) +
  geom_path() +
  geom_point(data=Prepared2.50 %>% filter(country == 'Ireland'),
             aes(x=Day2,y=Sum.Cases),
                 colour='red')+
  ylab('Total number of cases') + xlab('Days since fiftieth case') +
  ggtitle('Ireland COVID-19 in context',
          subtitle = 'Data and predictions') +
  geom_text_repel(data = Prepared2.50 %>%
                    filter(Day2 == max(Day2)),
                  aes(x=Day2, y=0.8*Sum.Cases, 
                      label=country),size=4) +
  scale_y_log10(labels=scales::comma) +
  scale_x_continuous() +
  theme_grey(base_size = 22) +
  theme(legend.position="none")

g

ggsave('images/Total_predictions_latest.png', g,
       units='cm', dpi=600, width=21, height=15)
```

## Graphs of latest daily predictions

```{r}
#Prediction <- Results %>%
#  mutate(Daily.Cases.Gom = Prediction.Gom - lag(Prediction.Gom)) %>% 
#  mutate(Daily.Cases.Lower = Lower.Gom - lag(Lower.Gom)) %>% 
#  mutate(Daily.Cases.Upper = Upper.Gom - lag(Upper.Gom)) %>%
#  mutate(Observed = Total.Cases) %>%
#  mutate(Predicted = Daily.Cases.Gom) %>%
#  mutate(Date = (Day -1) + ymd('2020-03-03')) 

#g <- ggplot(data = Prediction)  +
#  geom_point(aes(x = Date, y = Observed),colour='red') +
#  geom_line(aes(x = Date, y = Daily.Cases.Gom),colour='blue',alpha=0.4) #+
#  geom_ribbon(aes(x=Date, ymin=Daily.Cases.Lower, #ymax=Daily.Cases.Upper),
#            colour='blue', fill='lightblue', alpha=0.3) +
 # geom_point(aes(x = Date, y = Predicted),colour='blue',alpha=0.4) +
#  ylab('Daily reported cases') +
#  ggtitle('Reported daily cases of COVID-19 infection',
#          subtitle = 'Observed and predicted (Gompertz model)') +
#  theme_grey(base_size = 22)  +#
#  scale_x_date(date_breaks = "1 week", date_labels = "%d-%m")

#g

#ggsave('images/Daily_predictions_latest_Gom.png',g,units='cm',dpi=600,width=21,height=15)
```


```{r}
#Prediction <- Results %>%
#  mutate(Daily.Cases.LL = Prediction.LL - lag(Prediction.LL)) %>% 
#  mutate(Daily.Cases.Lower = Lower.LL - lag(Lower.LL)) %>% 
#  mutate(Daily.Cases.Upper = Upper.LL - lag(Upper.LL)) %>%
#  mutate(Observed = Total.Cases) %>%
#  mutate(Predicted = Daily.Cases.LL) %>%
#  mutate(Date = (Day -1) + ymd('2020-03-03')) #

#g <- ggplot(data = Prediction)  +
#  geom_point(aes(x = Date, y = Observed),colour='red') +
#  geom_line(aes(x = Date, y = Daily.Cases.LL),colour='blue',alpha=0.4) +
#  geom_ribbon(aes(x=Date, ymin=Daily.Cases.Lower, #ymax=Daily.Cases.Upper),
#            colour='blue', fill='lightblue', alpha=0.3) +
#  geom_point(aes(x = Date, y = Predicted),colour='blue',alpha=0.4) +
#  ylab('Daily reported cases') +
#  ggtitle('Reported daily cases of COVID-19 infection',
#          subtitle = 'Observed and predicted LL5 model') +
#  theme_grey(base_size = 22)  +
#  scale_x_date(date_breaks = "2 week", date_labels = "%d-%m")
#g

#ggsave('images/Daily_predictions_latest_LL.png',g,units='cm',dpi=600,wid#th=21,height=15)
```

```{r}
Ireland <- Ireland_fixed %>% filter(date>as_date('2020-03-03')) %>%
  mutate(casesSum7 = roll_sum(cases,n=7,fill=0,align='right')/7) %>%
  mutate(dateN = as.numeric(date))
ggplot(data=Ireland,aes(x=date,y=cases)) +geom_line()
ggplot(data=Ireland,aes(x=date,y=cases)) +geom_line() +scale_y_log10()
ggplot(data=Ireland,aes(x=date,y=casesSum7)) +geom_line()
ggplot(data=Ireland,aes(x=date,y=casesSum7)) +geom_line() +scale_y_log10()

Nadir0 <- as_date('2020-03-03')
Peak0 <- as_date('2020-04-27')
Gap0 <- Peak0-Nadir0 ; Gap0 # 55 days

Peak1 <- as_date('2020-04-27')
Nadir1 <- as_date('2020-06-26')
Gap1 <- Nadir1 - Peak1 ; Gap1 # 60 days

Nadir2 <- as_date('2020-07-06')
Peak2 <- as_date('2020-10-19')
as.numeric(Peak2)
Gap2 <- Peak2 - Nadir2; Gap2 # 105 days

coef(lm(data = Ireland %>% filter(date <= Peak0 & date >= Nadir0 ),
        cases ~ date)) # 12.8
12.9*55
coef(lm(data = Ireland %>% filter(date >= Peak1 & date <= Nadir1 ),
   cases ~ date)) # -5.4
60*-5.4
coef(lm(data = Ireland %>% filter(date <= Peak2 & date >= Nadir2 ),
   cases ~ date)) # 7.9
105*7.9

M0 <- lm(data = Ireland %>% filter(date <= Peak0 & date >= Nadir0 ),
   I(log10(cases)) ~ date)
M1 <- lm(data = Ireland %>% filter(date >= Peak1 & date <= Nadir1 ),
   I(log10(cases)) ~ date)
M2 <- lm(data = Ireland %>% filter(date <= Peak2 & date >= Nadir2 ),
   I(log10(cases)) ~ date)


summary(M0)
summary(M1)
summary(M2)

ggplot(data=Ireland,aes(x=date,y=casesSum7)) + geom_line() + scale_y_log10() + #geom_abline(slope=coef(M0)['date'],intercept=coef(M0)['(Intercept)'], color='blue') +
geom_abline(slope=coef(M1)['date'],intercept=coef(M1)['(Intercept)'], color='green') +
geom_abline(slope=coef(M2)['date'],intercept=coef(M2)['(Intercept)'], color='red') +
  ggtitle('What has happened') +
  ylab('7 day rolling average of cases') +
  xlab('Date') + 
  scale_x_date(date_breaks= "14 days", date_minor_breaks = "7 days",
               date_labels = "%d/%m", limits = as_date(c("2020-03-01", "2021-01-01"))) +
  geom_vline(xintercept = as_date('2020-12-25'), color='red', size=4) +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1)) 

  # y - y1 = m(x - x1) is the desired line
  # y1 is 3 m is -0.0183012 x1 is 18554
  # y - 3 = -0.0183012(x - 18554)
  # y = 3 -0.0183012(x - 18554)
  # y = -0.0183012*x + (3 + 185554*0.0183012)
  # y = 3398.861 - 0.0183012 * x
  # 16981498
  #342.6
  
  ggplot(data=Ireland,aes(x=date,y=log10(cases))) + geom_line() + 
    geom_abline(slope=coef(M1)['date'],intercept=coef(M1)['(Intercept)'], color='green') +
    geom_abline(slope=coef(M2)['date'],intercept=coef(M2)['(Intercept)'], color='red') +
    geom_abline(slope=coef(M1)['date'],intercept=556.8, color='blue') +
    ggtitle('What may happen') +
    ylim(0,4) +
    scale_x_date(date_breaks= "14 days", date_minor_breaks = "7 days",
                 date_labels = "%d/%m", limits = as_date(c("2020-03-01", "2021-03-30")))
```